### 2018-11-27 Writeup/PERT_Paper.tex
Reverse the change of Section 5

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{color}
\usepackage{amssymb}
\usepackage{accents}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage{framed}
\usepackage{natbib}
\usepackage{float, subfig}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage{amsthm}
\usepackage{grffile}
\usepackage{sidecap}
\usepackage{pbox}
\usepackage{algorithm}
\usepackage{longtable}
\usepackage[noend]{algpseudocode}

@@

	The management of complex projects through optimization has a rich history in operations research, beginning with the program evaluation and review technique (PERT) \cite{malcolm1959application} and the critical path method (CPM) of \cite{kelley1961criticalpath}; see S{\"o}derlund \cite{soderlund2004building} for an overview and see \cite{Elmaghraby77}. To formulate a mathematical model, a project is viewed as a collection of activities, each of which has some duration and may consume some resources. There are precedence relationships between activities due to logical or technological considerations. The objective is to complete all activities in minimum time. We represent such a project using an acyclic activity network, where the length of each link shows the duration of an activity, and an edge's predecessors capture the precedence relationships. An activity cannot start until all its predecessors are completed. Multiple activities can be processed at the same time without a upper limit, as long as the precedence requirement is satisfied. More details about the activity network are discussed in \cite{Elmaghraby77}.\\

@@

	In the planning stage of a project, decisions are made to crash a certain set of activities in order to minimize the project's timespan. Here crashing an activity means shortening its duration. In this paper, a discrete set of crashing options are available, each with fraction specifying the resulting reduction in the activity's duration. An activity can be crashed at most once and with a single option. Each option incurs a certain cost, and the total cost of crashing cannot exceed the budget. Optimization of such static crashing problems is introduced in \cite{fulkerson1961network, kelley1961criticalpath}. Under a finite set of crashing options, the problem can be modeled as a mixed integer program. The static model can be extended to incorporate uncertainty of activity durations. Monte Carlo simulation methods are used to estimate the expected project span given distributions of activity lengths, since it is is difficult to express analytically the expected project span \cite{burt1971conditional,van1963letter}. Heuristics and simulation-based algorithms have been developed to solve the stochastic project crashing problem \cite{aghaie2009ant,bowman1994stochastic,ke2014genetic,kim2007heuristic}. Another approach to handle the uncertainty is robust optimization, in which the objective is to minimize the worst case project span within a specified uncertainty set. While affinely adaptive recourse decisions are computationally tractable as linear, or second-order cone, programs, this restriction may lead to suboptimal solutions \cite{chen2008linear,cohen2007stochastic}. However, once recourse decisions can take general form, the robust model is only tractable with hypercube uncertainty sets, which is a trivial case because the worst-case scenario in the uncertainty set is always the point where every activity takes its upper bound \cite{wiesemann2012robust}. Ahipasaoglu et al. \cite{ahipasaoglu2016distributionally} propose a distributionally robust optimization scheme applied to a PERT network, which reformulates the problem as a semidefinite program or a copositive program, depending on the description of uncertainty.\\

@@

	The deterministic project crashing optimization problem has been known since the 1960s \cite{fulkerson1961network, kelley1961criticalpath}. Suppose the activity network is represented by a directed graph \(\mathcal{G} = (I,\mathcal{A})\), where the set of activities is denoted by \(I\) and their precedence relationship is characterized by \(\mathcal{A}\). The arc \((i,j) \in \mathcal{A}\) indicates that activity \(i\) has to be finished before the start of activity \(j\). We create two dummy activities \(S\) and \(T\) to represent the start and the end of entire project. \(S\) should precede every activity \(i \in I \backslash \{S\}\) and \(T\) should succeed every activity \(i \in I \backslash \{T\} \). \\
	For each activity \(i \in I\), the nominal length \(D_i\) is given. We can apply a finite set of crashing options, \(j \in J_i\), and one unit application of each option incurs a cost of \(b_{ij}\), while it decreases the length of activity \(i\) by \(D_ie_{ij}\). The total cost of crashing options cannot exceed a given budget, \(B\). We assume each activity can be crashed at most with one unit and the problem can be formulated as follows:
	In this formulation, \(t_i\) represents the starting time of activity \(i \in I\). We aim to minimize the project span, which is the starting time of the terminal activity, \(t_T\). Constraint~\eqref{cons:dSep} guarantees the precedence relationship: if activity \(i\) precedes activity \(k\), activity \(k\) cannot start until activity \(i\) is finished, since the finishing time of activity \(i\) is \(t_i + D_i (1- \sum_{j \in J_i} x_{ij}e_{ij})\). Constraint~\eqref{cons:dBudget} is the budget constraint and constraint~\eqref{cons:dSingleBudget} ensures that no more than one unit of crashing option can be applied to an activity. \\
	In our problem, we assume at most one stochastic disruption can occur at a random time in the project span. We also assume for each activity \(i \in I\), the crashing decision for that activity needs to be made at the start of that activity, and cannot be changed afterwards. The disruption does not affect the started activities but changes the length of those that have not started according to a random distribution. It is usually hard to compute the recourse function directly when parameters are distributed according to continuous distributions, and therefore SAA is proposed to approximate the problem by a large-scale deterministic optimization problem with a finite scenario set \cite{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a discrete finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, \(\xi^\omega\), consists of the timing of disruption, \(H\), and the changed duration of each activity, \(d_i, \forall i \in I\). \\

@@

\section{Experimental Results} \label{sec:results}
%	\textcolor{blue}{Test results:
%	\begin{itemize}
%		\item Describe the test examples: where do they come from, number of nodes, how are they generated.
%		\item Cite Sen's paper and describe the implementation.
%		\item Results.
%	\end{itemize}}
	In this section, we address the following questions with our computational results:
	\begin{enumerate}
		\item What is the value of model~\eqref{prob:extensive} that takes account of randomness in both timing and magnitude of a disruption? In other words, what are other alternative solutions that are easier to obtain but possibly inferior? How does quality of the solution to model~\eqref{prob:extensive} compare to those alternative solutions?
		\item How does the solution quality improve as the number of samples used in sample average approximation increases? 
		\item What is the time improvement of Algorithm~\ref{alg:Cut} compared to solving extensive formulation~\eqref{prob:extensive} using state-of-the-art MIP solvers? Other algorithms solving 2SMIP with mixed binary recourse?
	\end{enumerate}
	We first introduce the PERT network used for test and distribution information used to characterize uncertainty in Section~\ref{subsec:example}. In Section~\ref{subsec:value} we construct a series of deterministic and semi-deterministic alternatives to model~\eqref{prob:extensive}, perform out-of-sample tests, and present the better quality result of solution to model~\eqref{prob:extensive} compared to the alternative ones. We test the effect of different simulation budgets on solution quality and computational performance in Section~\ref{subsec:budget}. Finally in Section~\ref{subsec:time} we present the computational advantage of Algorithm~\ref{alg:Cut} compared to solving extensive formulation.\\
	%  and the ancestral Benders decomposition method (\textcolor{blue}{Cite Sen's two papers}.)
	\newline
	All tests are run on a server with 20 Intel Xeon cores at 3.1 GHz and 256 GB of RAM. All models are constructed using version 0.18.0 of the JuMP package \cite{DunningHuchetteLubin2017} on the Julia platform. All linear programs and mixed integer programs are solved by \textcolor{blue}{CPLEX} with both the feasibility tolerance and the integrality tolerance as \(10^{-7}\). We refer interested readers to \textcolor{blue}{GitHub repository} for the input data and the code used in this paper. %Gurobi 8.01 \cite{gurobi2016}

	\subsection{Test Cases Construction} \label{subsec:example}
%	\textcolor{blue}{Where we find our tests:
%		\begin{itemize}
%			\item Plambeck paper: Case 11
%			\item V42: Case 14
%			\item Case 19 from Elmgrabhy
%			\item Case 55 from Elmgrabhy
%			\item Randomized network 75, 100 nodes
%			\item Plambeck paper: Case 110
%	\end{itemize}}
	\noi Although it is easy to construct an activity network with randomly distributed disruption time and magnitude, not all instances constructed can well illustrate properties of our problem. Therefore, we construct our test cases based on the activity networks used in the previous literature, which have been shown of research value. We use two activity networks from \textcolor{blue}{Plambeck et al.}, one with 11 activities and the other with 110 activities. We also take two examples from Elmgrabhy \cite{Elmaghraby77}, one with 19 activities and the other with 55 activities. We manually create one activity network with 14 activities and generate two random networks with 75 and 100 activities, respectively using \textcolor{blue}{RanGen toolbox}. In the following section, we use ``Case \(X\)" to denote the test case with \(X\) activities.\\
	\newline
	For every test case, the timing of disruption follows a lognormal distribution since its value has to be nonnegative. The disruption magnitude of activities follows an exponential distribution of which the parameter differs among activities. The exponential distribution can best model the drastic change caused by a disruption.
	\subsection{Simulation Budget} \label{subsec:budget}
	We first examine the quality of solutions generated from different sizes of samples. It is well known that the objective value obtained by SAA approaches the objective function of the original problem as the sample size grows to infinity, and the expected value of the SAA objective value improves as the sample size increases~\cite{shapiro2009lectures}. However, when the sample size increases, the computational cost increases as well. Understanding how the solution quality changes according to different sample sizes is important to help us decide a proper sample size while achieving a decent solution. For each test case in Section~\ref{subsec:example}, we first obtain \(20\) batches of the optimal solution to model~\ref{prob:extensive} by Algorithm~\ref{alg:Cut}, with sample sizes \(n = 10, 50, 100, 200, 500\) and \(1000\). For each of those solutions, we test against a set of \(5000\) samples and obtain the object value. We compare the 
	\subsection{Value of a Fully Stochastic Model} \label{subsec:value}
	In this section, we compare the quality of five solutions obtained under different possible problem settings to show the value of modeling both the random timing and magnitude of the disruption. First, we can solve a deterministic project crashing problem assuming no disruption occurs (denoted as ``DET"). Three semi-stochastic alternatives can be constructed assuming: both timing and magnitude of the disruption are deterministic at their expected values (denoted as ``EXP"), the timing is random but the magnitude is at its expected value (denoted as ``\(H\)Only"), and the magnitude is random but the timing is at its expected value (denoted as ``\(d\)Only"). Finally, we construct the full stochastic model where both the timing and the magnitude are random (denoted as ``FULL"). 
	\subsection{Computational Performance} \label{subsec:time}
\bibliographystyle{plain}
\bibliography{PERT_Bib}

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{color}
\usepackage{amssymb}
\usepackage{accents}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage{framed}
\usepackage[square,sort,comma,numbers]{natbib}
\usepackage{url}
\usepackage{float, subfig}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage{amsthm}
\usepackage{grffile}
\usepackage{sidecap}
\usepackage{pbox}
\usepackage{algorithm}
\usepackage{longtable}
\usepackage[noend]{algpseudocode}

@@

	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath} for the deterministic project management problems; see S{\"o}derlund \cite{soderlund2004building} for an overview. A project is a collection of activities \(I\) of some duration \(D_i, i \in I\), and there are precedence relationships, \(\mathcal{A} \subseteq I \times I\), between activities due to logical or technological considerations. A project can be characterized as an acyclic activity network \(\mathcal{G} = (I,\mathcal{A})\). An arc \((i,j) \in \mathcal{A}\) indicates that activity \(i\) has to be finished before the start of activity \(j\), and its length shows the duration of an activity \(D_i\). An activity cannot start until all its predecessors are completed, while multiple activities can be processed at the same time without a upper limit, as long as the precedence requirement is satisfied. More details about the activity network are discussed in Elmaghraby~\cite{Elmaghraby77}. In this setting, ``crashing" is an action that consumes certain amount of resource and shorten the duration of activities accordingly~\cite{kuhl2008dynamic}. We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), while it decreases the length of activity \(i\) by \(D_ie_{ij}\); i.e., the crashed duration is proportional to both the nominal activity duration and the crashing decision. The total cost of crashing options cannot exceed a given budget, \(B\). The deterministic optimization problem for crashing decisions has been proposed since the 1960s \cite{fulkerson1961network, kelley1961criticalpath}. The goal is to complete all activities in minimum time by allocating resources under budget constraints. We can create two dummy activities \(S\) and \(T\) to represent the start and the end of entire project. \(S\) should precede every activity \(i \in I \backslash \{S\}\) and \(T\) should succeed every activity \(i \in I \backslash \{T\} \). Then the objective is then equivalent to minimize the starting time of activity \(T\). Thus, we formulate the deterministic project crashing optimization problem as:
		& t_i \geq 0  \qquad \qquad \forall \,i \in I \label{cons:dNonnegt}\\
		& 0 \leq x_{ij} \leq 1  \qquad \qquad \forall \,i \in I, j \in J_i. \label{cons:dxub}
	In this formulation, \(t_i\) represents the starting time of activity \(i \in I\). We aim to minimize the project span, which is the starting time of the terminal activity, \(t_T\). Constraint~\eqref{cons:dSep} guarantees the precedence relationship: if activity \(i\) precedes activity \(k\), activity \(k\) cannot start until activity \(i\) is finished, since the finishing time of activity \(i\) is \(t_i + D_i (1- \sum_{j \in J_i} x_{ij}e_{ij})\). Constraint~\eqref{cons:dBudget} is the budget constraint and constraint~\eqref{cons:dSingleBudget} ensures that no more than one unit of crashing option can be applied to an activity. In constraint~\eqref{cons:dxub}, we can put an artificial upper bound \(1\) on all crashing decisions \(x_{ij}\) by taking proper values of \(e_{ij}\). \\

@@

	The deterministic model can be extended to incorporate uncertainty of activity durations, starting with the program evaluation and review technique (PERT) \cite{malcolm1959application}. The PERT model assumes that activities are independently distributed with beta distributions, and the expected project span can be approximated by a normal distribution. However, this assumption may not be accurate for a general case~\cite{Elmaghraby77}. Therefore, assuming the activity durations follow a given probability distribution, Monte Carlo simulation methods can be used to estimate the expected project span, since it is is difficult to express analytically the expected project span \cite{burt1971conditional,van1963letter}. Heuristics and simulation-based algorithms have been developed to solve the stochastic project crashing problem \cite{aghaie2009ant, bowman1994stochastic, ke2014genetic, kim2007heuristic}. Another approach to handle the uncertainty is robust optimization, in which the objective is to minimize the worst case project span within a specified uncertainty set. While affinely adaptive recourse decisions are computationally tractable as linear, or second-order cone, programs, this restriction may lead to suboptimal solutions \cite{chen2008linear, cohen2007stochastic}. However, once recourse decisions can take general form, the robust model is only tractable with hypercube uncertainty sets, which is a trivial case because the worst-case scenario in the uncertainty set is always the point where every activity takes its upper bound \cite{wiesemann2012robust}. Ahipasaoglu et al. \cite{ahipasaoglu2016distributionally} propose a distributionally robust optimization scheme applied to a PERT network, which reformulates the problem as a semidefinite program or a copositive program, depending on the description of uncertainty. The project crashing optimization problem finds wide applications in project management~\cite{demeulemeester2006project, jaselskis1991allocation,tonchia2018industrial}, machine scheduling~\cite{blazewicz1983scheduling,hall1996machine}, health services scheduling~\cite{cardoen2010operating}, chemical processes~\cite{li2008process}, and digital circuit sizing~\cite{kim2007heuristic}.\\

@@

	In our problem, we assume at most one stochastic disruption can occur at a random time in the project span, since for the applications we are interested in (natural disasters, terrorist attacks, major market crashes, etc.), it is unlikely for two disruptions to occur in entire the time horizon. For example, it is unlikely for two major earthquakes to hit the same area twice within a month. It is reasonable to assume for each activity \(i \in I\), the crashing decision for that activity needs to be made at the start of that activity, and cannot be changed afterwards, especially when contracts are involved in commitment of resources~\cite{oberlender1993project}. The disruption does not affect the started activities but changes the length of those that have not started according to a random distribution. It is usually hard to compute the recourse function directly when parameters are distributed according to continuous distributions, and therefore SAA is proposed to approximate the problem by a large-scale deterministic optimization problem with a finite scenario set \cite{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a discrete finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, \(\xi^\omega\), consists of the timing of disruption, \(H\), and the changed duration of each activity, \(d_i, \forall i \in I\). \\

@@

\section{Experimental Results} \label{sec:results}
%	\textcolor{blue}{Test results:
%	\begin{itemize}
%		\item Describe the test examples: where do they come from, number of nodes, how are they generated.
%		\item Cite Sen's paper and describe the implementation.
%		\item Results.
%	\end{itemize}}
	In this section, we address the following questions with our computational results:
	\begin{enumerate}
		\item What is the value of model~\eqref{prob:extensive} that takes account of randomness in both timing and magnitude of a disruption? In other words, what are other alternative solutions that are easier to obtain but possibly inferior? How does quality of the solution to model~\eqref{prob:extensive} compare to those alternative solutions?
		\item How does the solution quality improve as the number of samples used in sample average approximation increases? 
		\item What is the time improvement of Algorithm~\ref{alg:Cut} compared to solving extensive formulation~\eqref{prob:extensive} using state-of-the-art MIP solvers? Other algorithms solving 2SMIP with mixed binary recourse?
	\end{enumerate}
	We first introduce the PERT network used for test and distribution information used to characterize uncertainty in Section~\ref{subsec:example}. In Section~\ref{subsec:value} we construct a series of deterministic and semi-deterministic alternatives to model~\eqref{prob:extensive}, perform out-of-sample tests, and present the better quality result of solution to model~\eqref{prob:extensive} compared to the alternative ones. We test the effect of different simulation budgets on solution quality and computational performance in Section~\ref{subsec:budget}. Finally in Section~\ref{subsec:time} we present the computational advantage of Algorithm~\ref{alg:Cut} compared to solving extensive formulation and the ancestral Benders decomposition method (\textcolor{blue}{Cite Sen's two papers}.)\\
	\newline
	All tests are run on a server with 20 Intel Xeon cores at 3.1 GHz and 256 GB of RAM. All models are constructed using version 0.18.0 of the JuMP package \cite{DunningHuchetteLubin2017} on the Julia platform. All linear programs and mixed integer programs are solved by Gurobi 8.01 \cite{gurobi2016} with default parameters.

	\subsection{Test Cases Construction} \label{subsec:example}
	%	\textcolor{blue}{Where we find our tests:
	%		\begin{itemize}
	%			\item Plambeck paper: Case 11
	%			\item V42: Case 14
	%			\item Case 19 from Elmgrabhy
	%			\item Case 55 from Elmgrabhy
	%			\item Randomized network 75, 100 nodes
	%			\item Plambeck paper: Case 110
	%	\end{itemize}}
	\noi Although it is easy to construct an activity network with randomly distributed disruption time and magnitude, not all instances constructed can well illustrate properties of our problem. Therefore, we construct our test cases based on the activity networks used in the previous literature, which have been shown of research value. We use two activity networks from \textcolor{blue}{Plambeck et al.}, one with 11 activities and the other with 110 activities. We also take two examples from Elmgrabhy \cite{Elmaghraby77}, one with 19 activities and the other with 55 activities. We manually create one activity network with 14 activities and generate two random networks with 75 and 100 activities, respectively using \textcolor{blue}{RanGen toolbox}. In the following section, we use ``Case \(X\)" to denote the test case with \(X\) activities.\\
	\newline
	For every test case, the timing of disruption follows a lognormal distribution since its value has to be nonnegative. The disruption magnitude of activities follows an exponential distribution of which the parameter differs among activities. The exponential distribution can best model the drastic change caused by a disruption.
	\subsection{Simulation Budget} \label{subsec:budget}
	We first examine the quality of solutions generated from different sizes of samples. It is well known that the objective value obtained by SAA approaches the objective function of the original problem as the sample size grows to infinity, and the expected value of the SAA objective value improves as the sample size increases~\cite{shapiro2009lectures}. However, when the sample size increases, the computational cost increases as well. Understanding how the solution quality changes according to different sample sizes is important to help us decide a proper sample size while achieving a decent solution. For each test case in Section~\ref{subsec:example}, we first obtain \(20\) batches of the optimal solution to model~\ref{prob:extensive} by Algorithm~\ref{alg:Cut}, with sample sizes \(n = 10, 50, 100, 200, 500\) and \(1000\). For each of those solutions, we test against a set of \(5000\) samples and obtain the object value. We compare the 
	\subsection{Value of a Fully Stochastic Model} \label{subsec:value}
	I	In this section, we compare the quality of five solutions obtained under different possible problem settings to show the value of modeling random timing and magnitude of the disruption. First, we can solve a deterministic project crashing problem assuming no disruption occurs (denoted as ``DET"). Three semi-stochastic alternatives can be constructed assuming: both timing and magnitude of the disruption are deterministic at their expected values (denoted as ``EXP"), the timing is random but the magnitude is at its expected value (denoted as ``\(H\)Only"), and the magnitude is random but the timing is at its expected value (denoted as ``\(d\)Only"). Finally, we construct the full stochastic model where both the timing and the magnitude are random (denoted as ``FULL"). 
	\subsection{Computational Performance} \label{subsec:time}
\bibliographystyle{plainnat}
\bibliography{PERT_Bib}
